---
title: "ARMA/VECM Analysis"
author: "Naim Sejtanic"
output:
  html_document: default
  pdf_document:
    latex_engine: xelatex
  word_document: default
---

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
library(plyr)
library(magrittr)
library(dplyr)
library(urca)
library(doBy)
library(zoo)
library(rvest)
library(rmarkdown)
library(tinytex)
library(forecast)
library(vars)
knitr::opts_chunk$set(echo = TRUE)
```



## ARMA/VECM Analysis

We will create an Ex Post Forecast of Equity Returns using an ARMA model and a VAR model, then compare the results. The data used is up until 2017 and is in the pv.csv file.

First, we estimate an AR(1) model of equity returns, $re_t$, with the sample period ending December 2015, and generate ex post forecasts from January to September 2016.

Create the equity returns and dividend.
```{r}
d0 <- read.csv("pv.csv")
equityreturn =  100 * diff(log(d0$P))
dividend = 100 * diff(log(d0$Real.Dividend))
d1 <- d0[2:length(d0$P),] %>%
  mutate(re = equityreturn, rd = dividend)
```

AR(1) model with the sample period ending December 2015, we get the estimated model $re_t = 0.3528 + 0.2839 re_{t-1}$

```{r}
d2.train <- subset(d1, d1$Date <= 2015.12)
ar1 = arima(d2.train$re, order = c(1,0,0))
summary(ar1)
```
Generate ex post forecasts from January to September 2016.
```{r}
ar1_expost = predict(ar1, n.ahead = 9)
autoplot(forecast(ar1))
ar1_expost
```


Next, we estimate a VAR(1) model of equity returns, $re_t$, and dividend returns, $rd_t$, with the sample period ending December 2015, and generate ex post forecasts from January to September 2016. 
Combine the training data
```{r}
vardata <- cbind(d2.train$re, d2.train$rd)
colnames(vardata) <- c("re", "rd")
```

VAR(1) model of equity returns and dividend returns. 
```{r}
var1 = vars::VAR(vardata, p = 1)
summary(var1)
```
We get the estimated VAR model $$re_t = 0.2379+ 0.2865 re_{t-1} +0.1045 rd_{t-1}$$,  $$rd_t = 0.0628 - 0.0077 re_{t-1} +0.5358 rd_{t-1}$$


Ex post forecasts from January to September 2016 using VAR(1) model.
```{R}
var1_expost <- predict(var1, n.ahead = 9)
var1_expost 
```


Now, we will compute the MSE and the RMSE for each in order to determine Which is the better forecasting model.
MSE and RMSE of the AR(1) model for equity returns. 
```{r}
d3.test <- subset(d1, d1$Date > 2015.12 & d1$Date <2016.10)
ar1_mse = mean((ar1_expost$pred - d3.test$re)^2)
ar1_mse
ar1_rmse =sqrt(ar1_mse)
ar1_rmse
```

MSE and RMSE of the VAR(1) model for equity returns. 
```{r}
var1_mse = mean((var1_expost$fcst$re-d3.test$re)^2)
var1_mse
var1_rmse = sqrt(var1_mse)
var1_rmse

```

Because the AR(1) model has smaller MSE, it is more accurate than VAR(1) model.




Next, we can estimate an AR(1) model of dividend returns $rd_t = \rho_0+\rho_1 rd_{t-1}+v_t$ and combine this model with the estimated model in the first equation to generate forecasts of equity returns from October to December 2016.
```{r}
d4 <- subset(d1, d1$Date <= 2016.09)
model = lm(re ~ rd, data = d4)
summary(model)
```
$\hat{re_t} = 0.3779 - 0.1912 \hat {rd_t} = 0.3779 - 0.1912 (\hat{\rho_0}+ \hat{\rho_1} rd_{t-1})$


AR(1)model $rd_t = \rho_0+\rho_1 rd_{t-1}+v_t$ using the time before Oct 2016 to predict Oct2016 to Dec2016.

```{r}
rd.fit =  arima(d4$rd, order = c(1,0,0))
summary(rd.fit)
rd_expost = predict(rd.fit, n.ahead = 3)
rd_expost
```

$\hat{re_t} = 0.3779 - 0.1912 \hat {rd_t}$, where $\hat{rd_t} = 0.1242 + 0.5383{rd_{t-1}}$.
Plug the predicted dividend returns into the model, we get the forecast of equity returns from October to December 2016. 

```{r}
d4.test = subset(d1, d1$Date <= 2016.12 & d1$Date >=2016.10)
rdpred = c(rd_expost$pred[[1]], rd_expost$pred[[2]], rd_expost$pred[[3]])
d4.test = d4.test[1:length(d4.test$P),] %>%
  mutate(rdpred = rdpred)
re_forecast = c(summary(model)$coefficients[1,'Estimate'] +
                  summary(model)$coefficients[2, 'Estimate']*d4.test$rdpred )
re_forecast
```

Now, we will estimate an AR(2) model of dividend returns $rd_t = \rho_0+\rho_1 rd_{t-1}+\rho_2 rd_{t-2} +v_t,$ and combine this model in place of the AR(1) model just above to generate forecasts of equity returns from October to December 2016. 

Estimate AR(2) model and use it to forecast dividend returns from October to December 2016. 
```{r}
rd.fit2 =  arima(d4$rd, order = c(2,0,0))
summary(rd.fit2)
rd_expost2 = predict(rd.fit2, n.ahead = 3)
rd_expost2
```
The estimated AR(2)model is $\hat{rd_t} = 0.122+ 0.4463 \hat{rd_{t-1}}+ 0.1710 \hat{rd_{t-2}}$.

Plugging the predicted dividend returns into the model, we get the forecast of equity returns from October to December 2016. 
```{r}
d4.test2 = subset(d1, d1$Date <= 2016.12 & d1$Date >=2016.10)
rdpred2 = c(rd_expost2$pred[[1]], rd_expost2$pred[[2]], rd_expost2$pred[[3]])
d4.test2 = d4.test2[1:length(d4.test2$P),] %>%
  mutate(rdpred2 = rdpred2)
re_forecast2 = c(summary(model)$coefficients[1,'Estimate'] +
                 summary(model)$coefficients[2, 'Estimate']*d4.test2$rdpred2 )
re_forecast2
```